---
title: "The Basic R Markdown"
author: "Diana Gwinn"
date: "2024-09-04"
output: html_document
---

# Data description

This data will show the salaries of each employee

- **Variable  1:** Rank- Professor's rank either assistant professor, associate professor, or professor
- **Variable 2:** Discipline- Type of department the professor works in, either applied or theoretical
- **Variable 3:** Yrs since PhD- Number of years since the professor has the obtained their PhD
- **Variable 4:** Yrs of service- Number of years the professor has served the department and/or university
- **Variable 5:** Sex- Professor's sex of either male or female
- **Variable 6:** Salary- Professor's nine-month salary (USD)

## Data wrangling

Please follow the steps to wrangle the data.

- **STEP 1:**  *Load the packages*
```{r packages, include=TRUE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
```

- **STEP 2:** *Set the working directory*
```{r directory, echo=FALSE}
getwd()
setwd("C:/Users/diana/OneDrive/Documents/Module 3")
```

- **STEP 3:** 
  - *Load the dataset* 
  - *Dismiss the first three rows*
```{r csv, echo=TRUE, results=FALSE}
Salaries <- read_csv("Salaries.csv")
```

- **STEP 4:** 
  - *Select() first*
  - *Then rename()*
  - *Finally filter()*
```{r select, echo=FALSE, results=FALSE, warning=FALSE, message=FALSE }
Salaries_new <- Salaries %>%
  select(rank,discipline,sex,salary) %>% #Select columns of “rank”, “discipline”, “sex”, “salary”
  filter(!is.na(salary)) %>% #Filter the rows where “salary” is not missing value.
  group_by(rank, discipline, sex) %>%  #Group based on the required columns (“rank”,”discipline”, “sex”) and have the tally count of the rows.
  summarise(count = n ()) #Get the tally count of the observation

```

- **STEP 5:** 
  - *Spread() for a wider table*
```{r wide, echo=FALSE, results=TRUE }
wide <- Salaries_new %>%
  spread(key = sex, value = count, convert = FALSE, drop = TRUE, sep = NULL, fill=FALSE)
```

- **STEP 6:** 
  - *Gather() for a longer table*
```{r long, echo=FALSE, results=TRUE  }
long <- wide %>%
  gather(key = "sex", value = "n", Female,Male, na.rm=TRUE, convert = FALSE, factor_key = FALSE)
View(long) 
```

